---
title: "Lab 4"
author: "Alexandria Quintana"
date: "10/5/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#**Question 1:** The number of points scored by each team in a tournament is normally distributed, with mean μ = 32 and standard deviation σ = 7.
##What is the probability of:


###Part 1: A team scoring 20 points or fewer? Plot the density function and shade the area. 
```{r}
x <- 20
curve(dnorm(x, mean =32, 7), from = 0, to = 46)
cord.x <- c(0, seq(0, 20, by = 0.01),20)
cord.y <- c(0,dnorm(seq(0,20, 0.01), mean = 32, sd = 7),0)
polygon(cord.x, cord.y, col='green')
polygon(cord.x, cord.y, density = 10)


```


###Part 2: A team scoring more than 35 points? Plot the density function and shade the area. 
```{r}
x <- dnorm(35, mean = 32, sd = 7)
curve(dnorm(x, mean = 32, 7), from = 0, to = 53)
cord.x <- c(35, seq(35,53, by = 0.01),53)
cord.y <- c(0, dnorm(seq(35,53, 0.01),mean = 32, sd = 7),0)
polygon(cord.x, cord.y, col='green')
```


###Part 3: A team scoring between 20 and 40 points? Plot the density function and shade the area.
```{r}
partthree <- diff(pnorm(c(20,40), mean = 32, sd = 32))
print(partthree)
curve(dnorm(x, mean = 32,7),from = 10, to = 53)
cord.x <- c(20, seq(20,40, by = 0.01),40)
cord.y <- c(0, dnorm(seq(20,40, 0.01), mean = 32, sd = 7),0)
polygon(cord.x, cord.y, col='green')

```

#**Question 2:** The number of comments per post on a social media site is exponentially distributed, with the average post receiving ten comments. What percentage of posts get:


###Part 1: Fewer than three comments? Plot the density function and shade the area.
```{r}
pexp(3, rate = 0.1, lower.tail = TRUE)
curve(dexp(x,0.1), from = 0, to = 150)
cord.x <- c(0, seq(0,3, 0.1),3)
cord.y <- c(0, dexp(seq(0,3, 0.1)),0)
polygon(cord.x, cord.y, col="blue")
```


###Part 2: More than 20 comments? Plot the density function and shade the area.
```{r}
pexp(20, rate = 0.1, lower.tail = FALSE)
curve(dexp(x,0.1), from = 0, to = 150)
cord.x <- c(20, seq(20, 150, 0.1),150)
cord.y <- c(0, dexp(seq(20,150, 0.1),0.1),0)
polygon(cord.x, cord.y, col="blue")
```


###Part 3: Between five and ten comments? Plot the density function and shade the area.
```{r}
pexp(50, rate = .10) - pexp(20, rate = .10)
curve(dexp(x,0.1), from = 0, to = 100)
cord.x <- c(5, seq(5, 10, 0.1),10)
cord.y <- c(0, dexp(seq(5,10, 0.1),0.1),0)
polygon(cord.x, cord.y, col="blue")
```


#**Question 3:** Basic raster creation and calculation


###Part 1: Create a first raster (30 rows ×30 columns) and assign values to the raster based on random values from a uniform distribution. Plot the raster and the histogram.
```{r}
if(!require(raster)) install.packages("raster")
library(raster)
if(!require(sp)) install.packages("sp")
library(sp)
rasone <- raster(nrow = 30, ncols = 30, xmn=0, xmx=10,ymn=0,ymx=10)
rasone
set.seed(10)
rasone[] <- runif(ncell(rasone))
plot(rasone)
hist(rasone)

```


###Part 2: Create a second raster (30 rows ×30 columns) and assign values to the raster based on random values from a normal distribution. Plot the raster and the histogram.
```{r}
set.seed(15)
rastwo <- raster(nrow = 30, ncols = 30, xmn=0, xmx=10,ymn=0,ymx=10)
rastwo
rastwo[] <- rnorm(ncell(rastwo))
plot(rastwo)
hist(rastwo)
```


###Part 3: Create a third raster (30 rows ×30 columns) and assign values to the raster based on cell-by-cell addition of the first raster and the second raster. Plot the raster and the histogram.
```{r}

rasthree <- rasone + rastwo
plot(rasthree)
hist(rasthree)

```


###Part 4: Calculate the mean value of the third raster. Reclassify the third raster into a binary image: 1 (cell value > mean value) and 0 (cell value <= mean value). Save the reclassified image to your working directory as a TIFF image.
```{r}
rasmean  <- cellStats(rasthree, "mean")
rasthree[rasthree > rasmean] <- 1
rasthree[rasthree < rasmean] <- 0
plot(rasthree)
writeRaster(rasthree, filename = "rasthree.tif", overwrite=TRUE)

```

