---
title: "Geog533 Lab 8"
author: "Alexandria Quintana"
output: 
  html_notebook:
    toc: TRUE
    toc_float: TRUE
---

Complete the following exercises in Chapter 8 (Introduction to Regression Analysis) of the textbook [R] pages 247-250. 

## Question 1

This is Exercise 1 in Chapter 8 of the Textbook [R].

A regression of weekly shopping trip frequency on annual income (data entered in thousands of dollars) is performed on data collected from 24 respondents. The results are summarized below:

* Intercept: 0.46
* Slope: 0.19

|          | Sum of squares  | Degrees of freedom |  Mean square   |       *F*       |
|----------|-----------------|--------------------|----------------|-----------------|
|Regression|                 |                    |                |                 |
| Residual |       1.7       |                    |                |                 |
| Total    |       2.3       |                    |                |                 |


### Question 1(a)
Fill in the blanks in the ANOVA table.

```{r}

#regression sum of squares
rss <- 2.3-1.7
rss
#Regression df = 1
#Residual df
rdf <- 24-2
rdf
#Regression Mean Square
RMS <- .6 / 1
RMS
#Residual Mean Square
ResMS <- 1.7/22
ResMS
#F
f <- .6 /.07
f

```


|          | Sum of squares  | Degrees of freedom |  Mean square   |       *F*       |
|----------|-----------------|--------------------|----------------|-----------------|
|Regression|  `r .6`         | `r 1`              | `r.6`          |`r 8.5`          |
| Residual |       1.7       | `r22`              | `r .07`        |                 |
| Total    |       2.3       |  `r 23`            |                |                 |


### Question 1(b)
What is the predicted number of weekly shopping trips for someone making $50,000/year?
```{r}
#y=a+bx
a <- .46
b <- .19
x <- 50000
y <- (a + b*(50000))
m <- y/1000
m


```


### Question 1(c)
In words, what is the meaning of the coefficient 0.19?
The meaning of the coefficient .19 is this is where the regression line crossess the y-axis(ordinate)


### Question 1(d)
Is the regression coefficient significantly different from zero? How do you know?
```{r}

rc <- qf(.95,1,22 )
rc
#yes because the value is 4.3
```



### Question 1(e)
What is the value of the correlation coefficient?
```{r}
r2 <- 2.3/.6
r2
```





## Question 2

This is Exercise 6 in Chapter 8 of the Textbook [R].

The following data are collected in an effort to determine whether snowfall is dependent upon elevation: 

| Snowfall (inches) | Elevation (feet) |
|-------------------|------------------|
|       36          |       400        |
|       78          |       800        |
|       11          |       200        |
|       45          |       675        |

Using R, show your work on exercises (a) through (g).

### Question 2(a)
Find the regression coefficients (the intercept and the slope coefficient).
```{r}
#a = mean(y) - b(xmean)
#b = r(sd(y)/sd(x))

snowfall <- c(36,78,11,45)
elevation <- c(400,800,200,675)
df <- data.frame(snowf)
m <- lm(snowfall ~ elevation)
summary(m)
anova(m)

#intercept: -7.63
#slope:0.09

```


### Question 2(b)
Estimate the standard error of the residuals about the regression line.
```{r}
#11.25
```


### Question 2(c)
Test the hypothesis that the regression coefficient associated with the independent variables is equal to zero. Also place a 95% confidence interval on the regression coefficient.
```{r}
#f-stat:16.17
qf(.95,1,2 )
#18.51
#therefore we accept
```


### Question 2(d)
Find the value of $r^2$.
```{r}
.88
```


### Question 2(e)
Make a table of the observed values, predicted values, and the residuals.
```{r}
#a = mean(y) - b(xmean)
#b = r(sd(y)/sd(x))
#a
df$outcome <- mean(elevation) -(.09*mean(snowfall))
#r
cor.test(x = snowfall, y = elevation)
#b
df$predictor <- (.94*(sd(elevation))/sd(snowfall))
#residuals
df$residuals <- c(253.3,126.61)
newdata <- data.frame(df$outcome,df$predictor, df$residuals)
View(newdata)
                                
```


### Question 2(f)
Prepare an analysis of variance table portraying the regression results.
```{r}
anova(m)
```


|          | Sum of squares  | Degrees of freedom |  Mean square   |       *F*       |
|----------|-----------------|--------------------|----------------|-----------------|
|Regression   `r2047.77`     |       r`1`         | `r2047.777`    |   `r 16.173 `   |
| Residual |    `r253.23`    |       `r2`         |`r126.61 `      |                 |
| Total    |  `r2300.23`     |                    |                |                 |

### Question 2(g)
Graph the data and the regression line.
```{r}
plot(elevation, snowfall)
abline(m, col="red")

```




## Question 3

This is Exercise 10 in Chapter 8 of the Textbook [R].

Use R and the [Milwaukee dataset](http://spatial.binghamton.edu/geog533/data/Milwaukee_Sales.csv) to:

### Problem
perform a regression using sales price as the dependent variable and lot size as the independent variable.

### Solution
```{r}
url = "http://spatial.binghamton.edu/geog533/data/Milwaukee_Sales.csv"
mdata <- read.csv("http://spatial.binghamton.edu/geog533/data/Milwaukee_Sales.csv", header = TRUE)

m <- lm(mdata$SalePrice ~ mdata$LotSize)
summary(m)
```



## Question 4

This is Exercise 11 in Chapter 8 of the Textbook [R].

Use R and the Hypothetical [UK Housing Prices dataset](http://spatial.binghamton.edu/geog533/data/UK_Housing.csv) to:

### Question 4(a) 
perform a regression using house price as the dependent variable, and number of bedrooms as the independent variable;

```{r}
mdata <- read.csv("http://spatial.binghamton.edu/geog533/data/UK_Housing.csv", header = TRUE)

m <- lm(mdata$price ~ mdata$bedrooms)
summary(m)

```


### Question 4(b) 
(b) repeat part (a), using the number of bathrooms as the independent variable, and comment on the results.
```{r}
m <- lm(mdata$price ~ mdata$bedrooms)
summary(m)
```


## Question 5
Download the [hometown.csv](http://spatial.binghamton.edu/geog533/data/hometown.csv), append your name and hometown coordinates to the csv, then use [leaflet](https://rstudio.github.io/leaflet/) to create a web map (10 pt.)

```{r}
if(!require(leaflet)) install.packages("leaflet")
mdata <- read.csv("hometown.csv", header = TRUE)
m <- leaflet()
m <- addTiles(m)
m <- addMarkers(m , lng = mdata$longitude, lat = mdata$latitude )
m



```

