---
title: "Geog533 Lab 10"
author: "Alexandria Quintana"
output: 
  html_notebook:
    toc: TRUE
    toc_float: TRUE
---

## Question 1
Use R package UScensus2010county to complete the following tasks:  (20 pt.)

### Question 1(a)
Plot a map of New York counties using the plot function.
```{r}

if(!require(UScensus2010)) install.packages("UScensus2010") 
library(UScensus2010)
if(!require(UScensus2010county)) install.county("osx") 
if(!require(UScensus2010tract)) install.tract("osx") 
library(UScensus2010county)
library(UScensus2010tract)
data(new_york.county10)
shp <- new_york.county10
plot(shp)

```

### Question 1(b)	
Plot a map of New York counties using the qtm function.
```{r}
library(UScensus2010county) 
data(new_york.county10)
shp <- new_york.county10
df <- shp@data
plot(shp)
if(!require(tmap)) install.packages("tmap")
library(tmap)
qtm(shp)
qtm(shp,fill = "NAME10") 
qtm(shp,fill = "P0010001") 
```


### Question 1(c)	
How many counties in New York State?
```{r}
#62
```

### Question 1(d)	
What’s the 3-digit fips code of Broome County?
```{r}
#007
```

### Question 1(e)	
Compute descriptive statistics of the population column (P0010001), including total, minimum, maximum, mean, median, and skewness. 
```{r}
pop <- df$P0010001 
summary(pop)
```

### Question 1(f)	
Create a histogram and a boxplot of the population.
```{r}
hist(pop)
boxplot(pop)

```


## Question 2
Use R package UScensus2010tract to complete the following tasks:    (20 pt.)
```{r}

```

### Question 2(a)	
Plot a map of New York census tracts using the plot function.
```{r}
if(!require(UScensus2010)) install.packages("UScensus2010") 
library(UScensus2010)
if(!require(UScensus2010county)) install.county("osx") 
if(!require(UScensus2010tract)) install.tract("osx") 
library(UScensus2010county)
library(UScensus2010tract)
data(new_york.tract10)
shp <- new_york.tract10
plot(shp)
```

### Question 2(b)
Compute the total population based on census tracts.
```{r}
library(UScensus2010county) 
data(new_york.tract10)
shp <- new_york.tract10
df <- shp@data
plot(shp)

popp <- df$P0010001 
sum(popp)
```

### Question 2(c)
Select all census tracts in Broome County and plot the map. 
```{r}
library(rgdal)
df <- shp@data

sel <- df$county == "007"

shp[sel,]
plot(shp[sel,])
```

### Question 2(d)
What’s the total population of Broome County?
```{r}
ex <- sel
b <- ex
pop <- b$P0010001
sum(pop)


```

### Question 2(e)
Create a histogram and a boxplot of population based on census tracts of Broome County.
```{r}
library(UScensus2010county) 
data(new_york.tract10)
shp <- new_york.tract10
df <- shp@data
pop <- df$P0010001 
hist(pop)
boxplot(pop)
```

### Question 2(f)
Select the first five columns of the shapefile of Broome County census tract; add a new population ratio column (= census tract population / county population); save the new shapefile to the hard drive. 
```{r}
head(shp)

df$newcol <- (pop/popp)
View(df)

```


## Question 3

Use R packages raster and ncdf4 to complete the following tasks:     (20 pt.)
```{r}
if(!require(ncdf4)) install.packages("ncdf4")
library(raster)
library(ncdf4)
```

### Question 3(a)		
Load the multi-band raster dataset “NDVI.nc” into R.
```{r}
N <- raster("/Users/Alexandria/Downloads/NDVI.nc/")
          
```

### Question 3(b)		
Get the basic information about the dataset, including the number of rows, columns, cells, and bands; spatial resolution, extent, bounding box, and projection.
```{r}
N
nrow(N) 
ncol(N)
```

### Question 3(c)	
Aggregate all bands to generate a mean NDVI raster and a maximum NDVI raster; save the two new raster datasets to the hard drive. 
```{r}

mean(N)
max(N)
plot(N)
ndvi2000 <- N[[1:12]]
aggregate(N)


```

### Question 3(d)			
Plot the maps of monthly NDVI of the year 2001
```{r, warning = FALSE}
N2 <- brick("/Users/Alexandria/Downloads/NDVI.nc/")
N2
ndvi2000 <- N2[[13:24]]
plot(ndvi2000, 13:24)

```

### Question 3(e)	
Create histograms of monthly NDVI of the year 2001.
```{r, warning = FALSE}
hist(ndvi2000)
```

### Question 3(f)			
Plot the NDVI map of July 2000; click any location with data on the map and retrieve the NDVI time series for all years; plot the NDVI time series of the selected location. 
```{r}
plot(N2, 1)
cell <- cellFromXY(N2, cbind(15, 51))
cell.ts <- N2[cell]
cell.ts
plot(as.vector(cell.ts), type = "b")
values <- N2[50,50]
plot(as.vector(values), type = "b")



plot(N2, 1)
values <- click(N2, n=1, xy = TRUE)
values <- click(N2, n=1, xy=FALSE)
plot(as.vector(values), type="b")
```

