---
title: "Geog533 Lab 7 - Correlation"
author: "Alexandria Quintana"
output: 
  html_notebook:
    toc: TRUE
    toc_float: TRUE
---

## Question 1
This is Exercise 1 in Chapter 7 of the Textbook [R].

### Problem 1a
Find the correlation coefficient, *r*, for the following sample data on income and education:
```{r}
library(knitr)
obs <- seq(1:5)
income <- c(30,28,52,40,35)
edu <- c(12,13,18,16,17)
df <- data.frame(obs,income,edu)
names(df) <- c("Observation","Income ($*1000)","Education (Years)")
kable(df)
```

Solution 1a
```{r}
colone <- c(30,28,52,40,35)
first <- df$colone

coltwo <- c(12,13,18,16,17)
second <- df$coltwo

df <- data.frame(colone, coltwo)
View(df)

#inividual steps to be able to find R
mx <- mean(colone)
mx

my <- mean(coltwo)
my

df$xd <- colone - mx

df$my <- coltwo - my

df$prod <- df$xd*df$my

a <- sum(df$prod)
a

sx <- sd(colone)
sx

sy <- sd(coltwo)
sy

b <- 4*sx*sy
b
#T o find R
r = a/b
r

```


### Problem 1b
Test the null hypothesis ρ = 0.

Solution 1b
```{r}
t <- r*sqrt(3)/sqrt(1-r^2)
t


#Critical T test
ct <- qt(0.975, 3)
ct

plot(df)
plot(colone, coltwo)


#Secondary test 
cor.test(x = colone, y = coltwo)

#we reject
```


### Problem 1c
Find Spearman’s rank correlation coefficient for these data.

Solution 1c
```{r}
cor.test(df$colone, df$coltwo,alternative = "less", method = "spearman")
```


### Problem 1d
Test whether the observed value of rs from part (c) is significantly different from zero.

Solution 1d
```{r}
#we accept
```



## Question 2
This is Exercise 3 in Chapter 7 of the Textbook [R].

### Problem
The distribution of the t-statistic for testing the significance of a correlation coefficient has n – 2 degrees of freedom. If the sample size is 36 and α = 0.05, what is the smallest absolute value a correlation coefficient must have to be significant? What if the sample size is 80?

### Solution
```{r}

 
#sample size of 36
x <- 2/sqrt(36)
x 

#sample size of 80
xx <- 2/sqrt(80)
xx


```



## Question 3
This is Exercise 4 in Chapter 7 of the Textbook [R].

### Problem
Find the correlation coefficient for the following data:

```{r}
library(knitr)
Obs <- seq(1:4)
X <- c(2,8,9,7)
Y <- c(6,6,10,4)
df <- data.frame(Obs,X,Y)
kable(df)
```

### Solution
```{r}
colone <- c(2,8,9,7)
first <- df$colone

coltwo <- c(6,6,10,4)
second <- df$coltwo

df <- data.frame(colone, coltwo)
View(df)

mx <- mean(colone)
mx

my <- mean(coltwo)
my

df$xd <- colone - mx

df$my <- coltwo - my

df$prod <- df$xd*df$my

a <- sum(df$prod)
a

sx <- sd(colone)
sx

sy <- sd(coltwo)
sy

b <- 3*sx*sy
b

r = a/b
r

t <- r*sqrt(2)/sqrt(1-r^2)
t



ct <- qt(0.975, 2)
ct

plot(df)
plot(colone, coltwo)



cor.test(x = colone, y = coltwo)

```



## Question 4
This is Exercise 6 in Chapter 7 of the Textbook [R].

### Problem
Find the correlation coefficient between median annual income in the United States and the number of horse races won by the leading jockey, for the period 1984–1995. Test the hypothesis that the true correlation coefficient is equal to zero. Interpret your results.

```{r}
year <- 1984:1994
income <- c(35165,35778,37027,37256,37512,37997,37343,36054,35593,35241,35486)
races <- c(399,469,429,450,474,598,364,430,433,410,317)
df <- data.frame(year,income,races)
names(df) <- c("Year","Median income","Number of races won by leading jockey")
kable(df)
```


### Solution
```{r}
colone <- c(35165,35778,37027,37256,37512,37997,37343,36054,35593,35241,35486)
first <- df$colone

coltwo <- c(399,469,429,450,474,598,364,430,433,410,317)
second <- df$coltwo

df <- data.frame(colone, coltwo)
View(df)

mx <- mean(colone)
mx

my <- mean(coltwo)
my

df$xd <- colone - mx

df$my <- coltwo - my

df$prod <- df$xd*df$my

a <- sum(df$prod)
a

sx <- sd(colone)
sx

sy <- sd(coltwo)
sy

b <- 10*sx*sy
b

r = a/b
r

t <- r*sqrt(9)/sqrt(1-r^2)
t



ct <- qt(0.975, 9)
ct

plot(df)
plot(colone, coltwo)



cor.test(x = colone, y = coltwo)
#we accept 
```
we reject 


## Question 5
This is Exercise 7 in Chapter 7 of the Textbook [R].

### Problem
For the following ranked data, find Spearman’s r, and then test the null hypothesis (using a Type I error probability of 0.10) that the true correlation is equal to zero.

```{r}
library(knitr)
obs <- 1:6
x <- c(1,2,5,6,11,12)
y <- c(8,4,12,3,10,7)
df <- data.frame(obs,x,y)
names(df) <- c("Observation","Rank of x","Rank of y")
kable(df)
```


### Solution
```{r}
#compare p value to .1, greater than accept 
colone <- c(1,2,5,6,11,12)
first <-  df$colone
coltwo <- c(8,4,12,3,10,7)
second <- df$coltwo
cor.test(df$colone, df$coltwo,alternative = "less", method = "spearman")
#we accept
```



## Question 6
This is Exercise 8 in Chapter 7 of the Textbook [R].

### Problem
Find Pearson’s r for the following data, and then test the null hypothesis that the correlation coefficient is equal to zero. Use a Type I error probability of 0.05.

```{r}
library(knitr)
obs <- 1:6
x <- c(3.2,2.4,1.6,8.3,7.2,5.1)
y <- c(6.2,7.3,8.1,2.6,6.3,4.3)
df <- data.frame(obs,x,y)
names(df) <- c("Observation","x","y")
kable(df)
```


### Solution
```{r}
colone <- c(3.2,2.4,1.6,8.3,7.2,5.1)
first <-  df$colone
coltwo <- c(6.2,7.3,8.1,2.6,6.3,4.3)
second <- df$coltwo
cor.test(df$colone, df$coltwo,alternative = "less", method = "pearson")
```
we accept

## Question 7
This is Exercise 9 in Chapter 7 of the Textbook [R].

### Problem
Using R and the Milwaukee dataset, find the correlation between number of bedrooms and lot size.

### Solution
```{r}
mdata <- read.csv("Milwaukee_sales.csv", header = TRUE)
rooms <- mdata$Bedrms
lotsize <- mdata$LotSize

cor.test(x = rooms, y = lotsize)
```


## Question 8
This is Exercise 10 in Chapter 7 of the Textbook [R].

### Problem
Using R and the Hypothetical UK Housing Prices dataset, find the correlation between floor area and number of bedrooms.

### Solution
```{r}
mdata <- read.csv("UK_Housing.csv", header = TRUE)
floor <- mdata$floorarea
br <- mdata$bedrooms

cor.test(x = floor, y = br)
```


## Question 9
Use the **cars** data frame in the **datasets** package to perform the following tasks: 

### Problem 9a
Plot a scatterplot for the data frame (x: speed, y: dist)

Solution 9a
```{r}
df <- cars
speed <- df$speed
dist <- df$dist
plot(speed, dist)

```


### Problem 9b
How many rows in the data frame?

Solution 9b
```{r}
nrow(df)
```


### Problem 9c
Calculate Pearson’s correlation coefficient using the equation below:
$$r=\frac{\sum_{n}^{i=1}(x_i-\bar{x})(y_i-\bar{y})}{(n-1)s_{x}s_{y}}$$

Solution 9c
```{r}
colone <- df$speed
first <- df$colone

coltwo <- df$dist
second <- df$coltwo

df <- data.frame(colone, coltwo)
View(df)

mx <- mean(colone)
mx

my <- mean(coltwo)
my

df$xd <- colone - mx

df$my <- coltwo - my

df$prod <- df$xd*df$my

a <- sum(df$prod)
a

sx <- sd(colone)
sx

sy <- sd(coltwo)
sy

b <- 49*sx*sy
b

r = a/b
r

t <- r*sqrt(48)/sqrt(1-r^2)
t



ct <- qt(0.975, 48)
ct

plot(df)
plot(colone, coltwo)



cor.test(x = colone, y = coltwo)
```


### Problem 9d
Use the cor.test() function to find Pearson’s correlation coefficient and compare it to the one from part (c)

Solution 9d
```{r}
cor.test(df$colone, df$coltwo,alternative = "less", method = "pearson")
```
They are the same

### Problem 9e
Calculate Spearman’s rank correlation coefficient using the equation below:
$$r_{S} = 1 - \frac{6\sum_{i=1}^{n}d_{i}^{2}}{n^3-n}$$

Solution 9e
```{r}


df$rankedone <- rank(colone)
df$rankedtwo <- rank(coltwo)
df$difference <- df$rankedone - df$rankedtwo
df$prod <- (df$difference)^2
l <- sum(df$prod)

m <- 1 - (l*6) / ((50^3) - 50)
m


```



### Problem 9f
Use the cor.test() function to find Spearman’s rank correlation coefficient and compare it to the one from part (e)

Solution 9f
```{r}
cor.test(df$colone, df$coltwo,alternative = "less", method = "spearman")
```
They are the same after comparing the two
